---
title: "Final Project"
author: "Kate Harkleroad & Daniel Finley"
date: "12/13/2020"
output: html_document
---
***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, error=FALSE, message=FALSE)
```

##### This document goes the code compiled to create a model that estimates the sale price of homes in Ames, IA. It goes through several sections and ends with generating a submission file. Sections: 
* Cleaning
* Data prep
* Data exploration
* Modeling

#### Loading the Libraries
```{r, warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(rpart)
library(rpart.plot)
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(xgboost)
library(corrplot)
library(reshape2)
```

#### Reading in and setting up the datasets
```{r, warning=FALSE, error=FALSE, message=FALSE}
train <- read_csv("train.csv")
test <- read_csv("test.csv")

#setting up train variables for binding
trainID <- train$Id
salePrice <- train %>%
  select(Id, SalePrice)

#setting up test variables for binding
testID <- test$Id
test$SalePrice <- NA

#combining the test and train datasets for cleaning data purposes.
test_train <- bind_rows(train, test)

```

***

# Cleaning

```{r}
#Remove large NA Variables
test_train <- test_train %>% select(-c("PoolQC", "MiscFeature", "Alley", "Fence", "Utilities"))

```

```{r}
#View NA's, don't worry about SalePrice NA's, these are from the test dataset
colSums(sapply(test_train,is.na))

```

***

# Data Prep


```{r}
# setting up the qualities vector to handle ordinal data
Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
```
     *  Ex	Excellent
     *  Gd	Good 
     *  TA	Average 
     *  Fa	Fair 
     *  Po	Poor 
     *  NA	None

### Fireplace variables

Within Fireplace Quality, there are 1420 NAs between test and train datasets. 

The number of NAs in FireplaceQu matches the number of houses with 0 fireplaces. This means that I can safely replace the NAs in FireplaceQu with 'no fireplace'. 

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$FireplaceQu[is.na(test_train$FireplaceQu)] <- 'None'

test_train$FireplaceQu<-as.integer(revalue(test_train$FireplaceQu, Qualities))
table(test_train$FireplaceQu)
```
```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train %>%
  filter(test_train$Fireplaces == 0)
# 1420/2919
```

### Basement

```{r}
test_train <- test_train %>% 
  mutate(BsmtExposure = ifelse(is.na(BsmtExposure), "noBsmt", BsmtExposure),
         BsmtFinished = ifelse(BsmtUnfSF == 0 & BsmtExposure != "noBsmt", 1, 
                                  ifelse(BsmtExposure == "noBsmt", 0, 0)))
```

***Looking at basement values with NAs***
```{r}
test_train %>% 
  select(contains("Bsmt")) %>% 
  filter(BsmtExposure == "noBsmt" & is.na(BsmtFinType1) & is.na(BsmtFinType2))
```
***Populating Basement NA values***
  Changing the remaining Na values to 0 for integer variables and "noBsmt" for character variables.

```{r}
test_train[test_train$BsmtExposure == "noBsmt" & is.na(test_train$BsmtFinType1) & is.na(test_train$BsmtFinType2),] <- test_train %>%
  filter(BsmtExposure == "noBsmt" & is.na(BsmtFinType1) & is.na(BsmtFinType2)) %>% 
  mutate(BsmtQual = "None",
         BsmtCond  = "None",
         BsmtFinType1 = "None",
         BsmtFinSF1 = 0,
         BsmtFinType2 = "None",
         BsmtFinSF2 = 0,
         BsmtUnfSF = 0,
         TotalBsmtSF = 0,
         BsmtFullBath = 0,
         BsmtHalfBath = 0,
         BsmtFinished = 0)
```

***Remaining Basement Na's***
  3 more basement variables contain NA values even though they contain basements. Will use most common data values to populate these 3 remaining variables. 
```{r}
#look at remaining Na's
NAcol <- which(colSums(is.na(test_train)) > 0)
sort(colSums(sapply(test_train[NAcol], is.na)), decreasing = TRUE)
```

```{r, warning=FALSE, error=FALSE, message=FALSE}
#Look at Basement Variables with remaining NA's
table(test_train$BsmtCond)
table(test_train$BsmtQual)
table(test_train$BsmtFinType2)
#look at NA value
test_train %>% select(Id, contains("Bsmt")) %>% 
  filter(is.na(BsmtFinType2))

#Populate BsmtCond NA's with most common value: TA
test_train$BsmtCond[is.na(test_train$BsmtCond)] <- factor("TA")

#Populate BsmtQual
test_train$BsmtQual[is.na(test_train$BsmtQual)] <- factor("TA")

#Populate BsmtFinType2 
test_train$BsmtFinType2 [is.na(test_train$BsmtFinType2)] <- factor("Unf")

```


***BsmtQual: Evaluates the height of the basement***

Make this variable ordinal

       Ex	Excellent (100+ inches)	
       Gd	Good (90-99 inches)
       TA	Typical (80-89 inches)
       Fa	Fair (70-79 inches)
       Po	Poor (<70 inches
       NA	No Basement

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$BsmtQual<-as.integer(revalue(test_train$BsmtQual, Qualities))
table(test_train$BsmtQual)
```

***BsmtCond: Evaluates the general condition of the basement***

Make this variable ordinal

       Ex	Excellent
       Gd	Good
       TA	Typical - slight dampness allowed
       Fa	Fair - dampness or some cracking or settling
       Po	Poor - Severe cracking, settling, or wetness
       NA	No Basement

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$BsmtCond<-as.integer(revalue(test_train$BsmtCond, Qualities))
table(test_train$BsmtCond)
```

***BsmtExposure: Refers to walkout or garden level walls***

Make this variable ordinal.

       Gd	Good Exposure
       Av	Average Exposure (split levels or foyers typically score average or above)	
       Mn	Mimimum Exposure
       No	No Exposure
       NA	No Basement
 
```{r, warning=FALSE, error=FALSE, message=FALSE}
Exposure <- c('noBsmt'=0, 'No'=1, 'Mn'=2, 'Av'=3, 'Gd'=4)

test_train$BsmtExposure<-as.integer(revalue(test_train$BsmtExposure, Exposure))
table(test_train$BsmtExposure)
```

***BsmtFinType1: Rating of basement finished area***

Make this variable ordinal.

       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement
		
```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$BsmtFinType1[is.na(test_train$BsmtFinType1)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)

test_train$BsmtFinType1<-as.integer(revalue(test_train$BsmtFinType1, FinType))
table(test_train$BsmtFinType1)
```

***BsmtFinType2: Rating of basement finished area (if multiple types)***

Make this variable ordinal with the FinType vector.

       GLQ	Good Living Quarters
       ALQ	Average Living Quarters
       BLQ	Below Average Living Quarters	
       Rec	Average Rec Room
       LwQ	Low Quality
       Unf	Unfinshed
       NA	No Basement

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$BsmtFinType2[is.na(test_train$BsmtFinType2)] <- 'None'
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)

test_train$BsmtFinType2<-as.integer(revalue(test_train$BsmtFinType2, FinType))
table(test_train$BsmtFinType2)
```



### Lot variables

3 variables. One with 1 NA, and 2 complete variables.

***LotFrontage: Linear feet of street connected to property***

486 NAs. The most reasonable imputation seems to take the median per neighborhood.

```{r}
ggplot(test_train[!is.na(test_train$LotFrontage),], aes(x=as.factor(Neighborhood), y=LotFrontage)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
for (i in 1:nrow(test_train)){
        if(is.na(test_train$LotFrontage[i])){
               test_train$LotFrontage[i] <- as.integer(median(test_train$LotFrontage[test_train$Neighborhood==test_train$Neighborhood[i]], na.rm=TRUE)) 
        }
}
```

**LotShape: General shape of property**

No NAs. Values seem ordinal (Regular=best)

       Reg	Regular	
       IR1	Slightly irregular
       IR2	Moderately Irregular
       IR3	Irregular

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$LotShape<-as.integer(revalue(test_train$LotShape, c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)))
table(test_train$LotShape)
sum(table(test_train$LotShape))
```

**LotConfig: Lot configuration**

No NAs. Factor this variable

       Inside	Inside lot
       Corner	Corner lot
       CulDSac	Cul-de-sac
       FR2	Frontage on 2 sides of property
       FR3	Frontage on 3 sides of property
       
```{r}
ggplot(test_train[!is.na(test_train$SalePrice),], aes(x=as.factor(LotConfig), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue')+
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..))
```

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$LotConfig <- as.factor(test_train$LotConfig)
table(test_train$LotConfig)
sum(table(test_train$LotConfig))
```

**Lot section probably is not going to be a good variable to use to predict SalesPrice **


### Garage NAs and variables

**Altogether, there are 5 variables related to garages**

Variables have 157 NAs between test and train. 

First of all, I am going to replace all 157 missing **GarageYrBlt: Year garage was built** values with the values in YearBuilt (this is similar to YearRemodAdd, which also defaults to YearBuilt if no remodeling or additions).
```{r}
#remaing NA's
NAcol <- which(colSums(is.na(test_train)) > 0)
sort(colSums(sapply(test_train[NAcol], is.na)), decreasing = TRUE)

#Looking at all garage variables when GarageType is null
test_train %>% 
  filter(is.na(GarageType)) %>%
  select(contains("garage"))
```
**GarageType: Garage location**

Factor these variables

       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage
```{r, warning=FALSE, error=FALSE, message=FALSE}
#Changing garage type from NA to No Garage
test_train$GarageType[is.na(test_train$GarageType)] <- "No Garage"
test_train$GarageType <- as.factor(test_train$GarageType)
table(test_train$GarageType)

```

**GarageYrBlt**
```{r}
test_train$GarageYrBlt[is.na(test_train$GarageYrBlt)] <- test_train$YearBuilt[is.na(test_train$GarageYrBlt)]
```

From looking at the summary statistics, the max GarageYrBlt is 2207. Need to look in to this and change the GarageYrBlt to the value of the YearBuilt

```{r}
test_train %>% 
  filter(GarageYrBlt > 2020) %>% 
  select(Id, YearBuilt, GarageYrBlt)

#Changing GarageYrBlt to the value of YearBuilt
test_train$GarageYrBlt[test_train$Id == 2593] <- 2006

```

**GarageFinish: Interior finish of the garage**

The values are ordinal.

       Fin	Finished
       RFn	Rough Finished	
       Unf	Unfinished
       NA	No Garage       
```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$GarageFinish[is.na(test_train$GarageFinish)] <- 'None'
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)

test_train$GarageFinish<-as.integer(revalue(test_train$GarageFinish, Finish))
table(test_train$GarageFinish)
```


**GarageQual: Garage quality**

The values are ordinal.

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage
       
```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$GarageQual[is.na(test_train$GarageQual)] <- 'None'
test_train$GarageQual<-as.integer(revalue(test_train$GarageQual, Qualities))
table(test_train$GarageQual)
```

**GarageCond: Garage condition**

The values are ordinal.

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       NA	No Garage

```{r, warning=FALSE, error=FALSE, message=FALSE}
test_train$GarageCond[is.na(test_train$GarageCond)] <- 'None'
test_train$GarageCond<-as.integer(revalue(test_train$GarageCond, Qualities))
table(test_train$GarageCond)
```

***GarageType: Garage location***
		
       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage: 
       
```{r}
test_train <- test_train %>% 
  mutate(GarageType = ifelse(is.na(GarageType), "noGarage", GarageType))
```

***GarageArea: Size of garage in square feet + GarageCar: Size of garage in car capacity***

1 Na value for GarageArea and 1 NA for GarageCars. Both NA's on same record and Does not appear to have a garage, so setting sqft garagecars to 0


```{r}
test_train %>% select(Id, GarageArea, GarageType, GarageCond, GarageFinish, GarageQual) %>% 
  filter(is.na(GarageArea))

test_train$GarageArea[is.na(test_train$GarageArea)] <- 0

test_train %>% select(Id, GarageArea, GarageType, GarageCond, GarageFinish, GarageQual, GarageCars) %>% 
  filter(is.na(GarageCars))

test_train$GarageCars[is.na(test_train$GarageCars)] <- 0



```


### Masonry variables

**Masonry veneer type, and masonry veneer area**

Masonry veneer type has 8 NAs. Masonry veneer area has 8 NAs. If a house has a veneer area, it should also have a masonry veneer type. 


**Masonry veneer type**

Will check the ordinality below.

       BrkCmn	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       None	None
       Stone	Stone

```{r}
test_train$MasVnrType[is.na(test_train$MasVnrType)] <- 'None'
```

There seems to be a significant difference between "common brick/none" and the other types. I assume that simple stones and for instance wooden houses are just cheaper. I will make the ordinality accordingly. 

```{r, warning=FALSE, error=FALSE, message=FALSE}
Masonry <- c('None'=0, 'BrkCmn'=0, 'BrkFace'=1, 'Stone'=2)
test_train$MasVnrType<-as.integer(revalue(test_train$MasVnrType, Masonry))
table(test_train$MasVnrType)
```

**MasVnrArea: Masonry veneer area in square feet**

An integer variable.
```{r}
test_train$MasVnrArea[is.na(test_train$MasVnrArea)] <-0
```



### Electrical system

**Electrical: Electrical system**

1 NA. Values are categorical.

       SBrkr	Standard Circuit Breakers & Romex
       FuseA	Fuse Box over 60 AMP and all Romex wiring (Average)	
       FuseF	60 AMP Fuse Box and mostly Romex wiring (Fair)
       FuseP	60 AMP Fuse Box and mostly knob & tube wiring (poor)
       Mix	Mixed

```{r, warning=FALSE, error=FALSE, message=FALSE}
#imputing mode
test_train$Electrical[is.na(test_train$Electrical)] <- names(sort(-table(test_train$Electrical)))[1]

test_train$Electrical <- as.factor(test_train$Electrical)
table(test_train$Electrical)
sum(table(test_train$Electrical))
```

***Functional***
Functional: Home functionality (Assume typical unless deductions are warranted) 
2 NA's going to assume they are Typ(Typical Functionality)

```{r, warning=FALSE, error=FALSE, message=FALSE}
summary(factor(test_train$Functional))

test_train$Functional[is.na(test_train$Functional)] <- "Typ"

```
### Zoning
***MSZoning***
MSZoning: Identifies the general zoning classification of the sale.

4NA values 

```{r}
summary(factor(test_train$MSZoning))

test_train %>% 
  filter(is.na(MSZoning)) %>%
  select(Id, MSZoning, Neighborhood)
```
Creating a tree function to predict MSZoning values for Id = 1916, 2217, 2251, 2905
```{r}
tree.MSZoning <- rpart(MSZoning ~ ., data = test_train)

test_train[is.na(test_train$MSZoning), ]$MSZoning <- predict(tree.MSZoning,
                                                   newdata = test_train[is.na(test_train$MSZoning), ],
                                             type = "class")
```

Look at the MSZoning Values for Id = 1916, 2217, 2251, 2905
```{r}
test_train %>% 
  filter(Id == 1916 | Id == 2217 | Id == 2251 | Id == 2905) %>%
  select(Id, MSZoning, Neighborhood)
```

### SaleType

***SaleType: Type of sale***
1 NA
```{r, warning=FALSE, error=FALSE, message=FALSE}
summary(factor(test_train$SaleType))

test_train %>% 
  filter(is.na(SaleType)) %>%
  select(Id, SaleType, SaleCondition)

#Setting Sale Type to WD - the most common Sale Type
test_train$SaleType[is.na(test_train$SaleType)] <- "WD"


```

### Kitchen

***Kitchen Quality***
KitchenQual: Kitchen quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor
       
  1 NA value
  
```{r, warning=FALSE, error=FALSE, message=FALSE}
summary(factor(test_train$KitchenQual))

test_train %>% 
  filter(is.na(KitchenQual)) %>%
  select(Id, KitchenQual, KitchenAbvGr, OverallQual, OverallCond)

#Giving KitchenQual value TA (The most common value)

test_train$KitchenQual[is.na(test_train$KitchenQual)] <- "TA"

#Categorical Variable to Integer
test_train$KitchenQual<-as.integer(revalue(test_train$KitchenQual, Qualities))
table(test_train$KitchenQual)
```

### Exterior 

***Exterior1st & Exterior2nd***
Exterior1st: Exterior covering on house

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast	
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles
	
Exterior2nd: Exterior covering on house (if more than one material)

       AsbShng	Asbestos Shingles
       AsphShn	Asphalt Shingles
       BrkComm	Brick Common
       BrkFace	Brick Face
       CBlock	Cinder Block
       CemntBd	Cement Board
       HdBoard	Hard Board
       ImStucc	Imitation Stucco
       MetalSd	Metal Siding
       Other	Other
       Plywood	Plywood
       PreCast	PreCast
       Stone	Stone
       Stucco	Stucco
       VinylSd	Vinyl Siding
       Wd Sdng	Wood Siding
       WdShing	Wood Shingles
       
  1 NA on both Exterior1st and Exterior2nd
  
  
```{r, warning=FALSE, error=FALSE, message=FALSE}
summary(factor(test_train$Exterior1st))
summary(factor(test_train$Exterior2nd))

test_train %>% 
  filter(is.na(Exterior1st)) %>%
  select(Id, Exterior1st, Exterior2nd)

#NA's for Exterior1st and Exterior2nd are on the same record
#Setting values to the most common values

test_train$Exterior1st[is.na(test_train$Exterior1st)] <- "VinylSd"
test_train$Exterior2nd[is.na(test_train$Exterior2nd)] <- "VinylSd"


```
Exterior Quality
```{r}
test_train$ExterQual<-as.integer(revalue(test_train$ExterQual, Qualities))
table(test_train$ExterQual)
```
### Street
```{r}
#Street
test_train$Street <- ifelse(test_train$Street=="Pave", 1, 0)

```

	

```{r}
colSums(sapply(test_train,is.na))

```

### Data Prep Summary: 
That takes care of all of the NAs. We have also manipulated a lot of the data to make it easier to use. This next section will take a deeper look at the data we have prepped. 



# Data Exploration

**1. Take a look at the OverallQual Variable**
```{r}
test_train %>% 
  ggplot(aes(x = OverallQual, y = SalePrice)) +
  geom_point() +
  geom_smooth() +
  xlab("Overall Quality")
```


**2. Neighborhood**
```{r}
#table(test_train$Neighborhood)

ggplot(test_train[!is.na(test_train$SalePrice),], aes(x=Neighborhood, y=SalePrice)) +
        geom_bar(stat='summary', fun = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #->median sales price

#binning
ggplot(test_train[!is.na(test_train$SalePrice),], aes(x=reorder(Neighborhood, SalePrice, FUN=mean), y=SalePrice)) +
        geom_bar(stat='summary', fun = "mean", fill='blue') + labs(x='Neighborhood', y="Mean SalePrice") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #->median sales price
```
*The red dotted line shows the overall median price and the blue bars illustrate the median price per neighborhood. You can see here that there are a few neighborhoods where you can get a much high or lower sales price.*

**3. GrLivArea**
```{r}
#table(test_train$GrLivArea)
test_train %>% 
  ggplot(aes(x = GrLivArea, y = SalePrice)) +
  geom_point() +
  geom_smooth() +
  xlab("Above Ground Living Area")
```

**4. 1stFlrSF**
```{r}
test_train %>% 
  ggplot(aes(x = `1stFlrSF`, y = SalePrice)) +
  geom_point() +
  geom_smooth() +
  xlab("First Floor Sqft")
```

*There is an outlier here we may want to remove. *

**5. TotalBsmtSF**
```{r}
#table(test_train$TotalBsmtSF)
test_train %>% 
  ggplot(aes(x = `TotalBsmtSF`, y = SalePrice)) +
  geom_point() +
  geom_smooth() +
  xlab("Total Basement Sqft")
```

*Looks like possibly the same outlier?*


**6.Bathrooms**

*Combining bathroom variables and looking at their relationship to SalePrice*

```{r}
test_train <- test_train %>% 
  mutate(FullBath = as.numeric(FullBath),
         HalfBath = as.numeric(HalfBath * .5),
         BsmtFullBath = as.numeric(BsmtFullBath),
         BsmtHalfBath = as.numeric(BsmtHalfBath * .5),
         TotalBaths = as.numeric(FullBath + BsmtFullBath + HalfBath + BsmtHalfBath))
```

*Plotting new Total Bathroom variable*

```{r}
  ggplot(data = test_train, aes(x = TotalBaths, y = SalePrice)) +
  geom_point() + geom_smooth()
```

*There is that pesky outlier again*

**7. Total Home SQFT**

```{r}
#TotalBsmtSF, 1stFlrSF, 2ndFlrSF

test_train <- test_train %>%
  mutate(TotalSqft = TotalBsmtSF + `1stFlrSF` + `2ndFlrSF`)

#Plot Total Square Feet against SalePrice

ggplot(data = test_train, aes(x = TotalSqft, y = SalePrice)) + geom_point(col='black') + geom_smooth() + scale_y_continuous(breaks = seq(0, 800000, by = 100000), labels = comma)

#Plot 2
ggplot(data = test_train, aes(x = TotalSqft, y = SalePrice)) + geom_point(col='black') + geom_smooth(method = lm) + scale_y_continuous(breaks = seq(0, 800000, by = 100000), labels = comma)


```


 *Outlier IDs: 524 and 1299. *


**8. House age**
```{r, fig.width= 7, fig.height=10}
test_train$Age <- as.numeric(test_train$YrSold)-test_train$YearRemodAdd


h1 <- ggplot(data=test_train[!is.na(test_train$SalePrice),], aes(x=Age, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)

#New houses
test_train$IsNew <- ifelse(test_train$YrSold==test_train$YearBuilt, 1, 0)
table(test_train$IsNew)

hnew <- ggplot(test_train[!is.na(test_train$SalePrice),], aes(x=as.factor(IsNew), y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=6) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        theme_grey(base_size = 18) +
        geom_hline(yintercept=163000, linetype="dashed") #dashed line is median SalePrice

grid.arrange(h1, hnew)
```

*Downward trend in house age as to be expected. *


**Some additional plots**
```{r}
#Plot Age vs SalePrice
ggplot(data = test_train, aes(x = Age, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot Kitchen Qual vs SalePrice
ggplot(data = test_train, aes(x = KitchenQual, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot 1st Floor Sqft vs SalePrice
ggplot(data = test_train, aes(x = `1stFlrSF`, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot Basement Qual vs Sale Price
ggplot(data = test_train, aes(x = BsmtQual, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot Garage Cars vs Sale Price
ggplot(data = test_train, aes(x = GarageCars, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot total rooms above ground
ggplot(data = test_train, aes(x = TotRmsAbvGrd, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

#Plot Exterior Quality vs Sale Price
ggplot(data = test_train, aes(x = ExterQual, y = SalePrice)) + geom_point() + geom_smooth(method = lm)

ggplot(data = test_train, aes(x = Street, y = SalePrice)) + geom_point() + geom_smooth(method = lm)
```



*Remove outliers*
```{r}
test_train <- test_train[-c(524, 1299),]
```

```{r}
(saleprice_tree <- rpart(SalePrice~. , data = test_train))

rpart.plot(saleprice_tree)
```

**Our current top contenders: OverallQual, Neighboorhood, 1stFlrSF, GrLivArea, BsmtFinSF1 **



### Exploring Correlation: Dropping highlly correlated values: 

```{r, fig.height=7, fig.width=7}
numericVars <- which(sapply(test_train, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')

all_numVar <- test_train[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt", tl.cex = 0.7,cl.cex = .7, number.cex=.7)
```

*If two variables are highly correlated, drop the variable with a lower correlation with SalePrice: *

* GarageArea + GarageCars = drop GarageArea (0.62)
* GrLivArea + TotalSqFeet = drop GrLivArea
* 1stFlrSF + TotalBsmtArea = tie
* GarageYrBlt + Year Built = drop GarageYrBlt
* YearRemodAdd + Age = d YearRemodAdd
* TotalSqFeet + TotalBsmtSF = drop TotalBsmtSF
* total area and rooms above = drop TotRmsAbvGrd

```{r}
dropVars <- c('GarageArea', 'GrLivArea', 'GarageYrBlt', 'YearRemodAdd', 'TotalBsmtSF' )

test_train <- test_train[,!(names(test_train) %in% dropVars)]
```

# Modeling

***Splitting the data back in to test and train datasets***
```{r}
train <- test_train %>% 
  filter(Id %in% trainID) 
test <- test_train %>% 
  filter(Id %in% testID) %>% 
  select(-SalePrice)
```

# Looking at the distrubtion of Sales Price
```{r}
ggplot(data = train, aes(x= SalePrice)) + geom_histogram()
```

*Sale Price distribution slightly skewed will log sale price in model*


**Setting up the linear regression model and removing the Id variable from the train set**
```{r}
train <- train%>%
  select(-Id)

lm_model_1 <- lm(log(SalePrice) ~. , data = train)

summary(lm_model_1)
```

***Testing with limited variables***
```{r}

lm_model_2 <- lm(log(SalePrice) ~ Age+KitchenQual+
                   BsmtQual+TotRmsAbvGrd+
                   ExterQual+TotalSqft+
                   YearBuilt+TotalBaths+OverallQual, data = train)
summary(lm_model_2)


#RMSE
prediction <- predict(lm_model_2, train, type="response")
model_output <- cbind(train, prediction)

model_output$log_prediction <- log(model_output$prediction)
model_output$log_SalePrice <- log(model_output$SalePrice)

#Test with RMSE

#rmse(model_output$log_SalePrice,model_output$log_prediction)
```
Testing with Random forest
```{r}
#model_1 <- randomForest(SalePrice ~ ., data=train)


# Predict using the test set
#prediction <- predict(model_1, test)
#model_output <- cbind(test, prediction)


#model_output$log_prediction <- log(model_output$prediction)
#model_output$log_SalePrice <- log(model_output$SalePrice)

#Test with RMSE

#rmse(model_output$log_SalePrice,model_output$log_prediction)
```


```{r}
#RMSE
rmse <- function(actual, fitted){
  sqrt(mean((actual - fitted)^2))
}

rmse(train$SalePrice, fitted(lm_model_1))
# need to exponentiate the log value
rmse(train$SalePrice, exp(fitted(lm_model_1)))
```


Prepare the prediction data
```{r}
predicted_prices <- exp(predict(lm_model_1, newdata = test))
predicted_prices2 <- exp(predict(lm_model_2, newdata = test))

#Quick Look
head(predicted_prices)
head(predicted_prices2)
```

Prepare Submission File
```{r}
# create a dataframe with our results
our_submission <- tibble('Id' = as.integer(test$Id), 'SalePrice' = predicted_prices)
our_submission2 <- tibble('Id' = as.integer(test$Id), 'SalePrice' = predicted_prices2)

# save our file
write_csv(our_submission2, 'submission7.csv')

head(our_submission)
summary(our_submission)
head(our_submission2)

```
```

